Discrete Quantum Theories and Computing

### 2019-01-20 dissertation.tex
dissertation.lyx Exported to HTML, and use Microsoft Word to grammar check...

\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\frameL}[1][]{\ensuremath{f^{L{#1}}}}
\newcommand{\frameR}[1][]{\ensuremath{f^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }

@@

\item Hao-Chun Lee (李浩君) and Hsien-Ching Kao (高憲慶) for discussing
what I did in the prospective of a general PhD. in Physics, and working
later in more computational industry.

@@

Although it is convenient to assume physical quantities is as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick are made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
actually be an integer multiple of the elementary charge. A prefect
square cannot be draw physically. Even we don't draw it, but merely
ask whether it exists in the space. This question is still beyond
our ability to answer because whether the space itself is ultimately
discrete or continuous is still an unsolved question. Even if a prefect
square exists physically, we may not able to decide whether the ``square''
we discussed is really prefect or almost prefect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision in reality is limited generally three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Hence, when the quantities used to branch the computation closes to
zero, the branch chosen by a BCSS machine cannot reliably predict
the branch chosen by a realistic analog computer. Since two branches
of computations may not have any relations, the follow-up computational
paths may be significantly different, and these difference is unlikely
to be compensated by error analysis techniques easily like the butterfly
effect. Therefore, the computational power of analog computers might
be far weaker than the prediction based on this kind of theoretical
models.%
\begin{comment}
\todo{Maybe extended to consider probabilistic model? }
\end{comment}

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the Physical device? & 
Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Similar to its classical counterparts, quantum circuit model,
the most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states. To address the similar inconsistency, we first review the
``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}, where the probability amplitudes are in the field of complex numbers.
If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, taking into account the
precision of quantities could still result in a better model. However,
comparing to the precision of probability amplitudes, it is easier
to consider the precision of their inducing probabilities because
the idea of ``imprecise probabilities'' have well-studied classically
\cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }This idea will be extended
to quantum domain as quantum interval-valued probability measures
(QIVPM).

@@

Finally, in Section~\ref{discretequantumcomputingII}, we apply our
discrete quantum theory to the study of two representative algorithms,
the deterministic Deutsch-Jozsa algorithm and the probabilistic Grover
algorithm~\cite{NCbook,Mermin}. The first algorithm highlights the
role played by the size of the field $p$ in determining the actual
resources required for computation as the number of input bits $n$
increases, a concept nonexistent in conventional quantum computing.
The second algorithm highlights, in addition, the dependence of the
precision of measurement (via cardinal probabilities) on the size
of the field, another nonexistent concept in conventional quantum
computing.
\end{comment}
\begin{comment}
Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\end{comment}
Since many important quantum algorithms are probabilistic, we want
to find a notion of probability for quantum theories over finite fields.
The cardinal probability defined in Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
restricts states in some local regions of the vector spaces, where
a notion of inner product and probability could be recovered. In Sec.~\ref{discretequantumcomputingII},
this discrete quantum theory is applied on two representative algorithms,
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\begin{comment}
When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can actually be deduced from a set of abstract definitions
and axioms according to Gleason's theorem. Although we might hope
to deduce a discrete Born rule directly from a similar set of definitions
and axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability, and extend it with the conventional quantum theory.
\end{comment}
Despite the great success of cardinal probabilities, it is difficult
to define their arithmetic operations and expectation values. Since
they could be defined easily with real-valued probabilities in CQT,
we attempt to consider real-valued quantum probability measures over
finite fields (QPMFF) in Sec.~\ref{sec:Toward-IVPM}. In contrast
to CQT, our Gleason-like conditions over finite fields cannot deduce
a discrete Born rule in general. This impossible result suggests us
that the discrete and finite state spaces might not be compatible
with infinitely precise real-valued probability. Thus, no matter whether
the probability amplitudes are ultimately discrete or continuous,
we should study imprecise quantum probabilities formulated by quantum
interval-valued probability measures.

\begin{comment}
In the classical setting, there are several proposals for ``imprecise
probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\bar{\mu}(E)$ of an event~$E$ is generalized
from a single \emph{real number} to an \emph{interval}~$[l,r]$,
where~$l$ intuitively corresponds to the strength of evidence for
the event~$E$ and~$1-r$ corresponds to the strength of evidence
against the same event. Given a sample space~$\Omega$ and a set
of intervals~$\mathscr{I}$, similar to a classical probability measure
$\mu:2^{\Omega}\rightarrow\left[0,1\right]$, a classical interval-valued
probability measure (IVPM) $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
needs to satisfy some coherent axioms. By satisfying the convexity
axiom \cite{Shapley1971,GilboaSchmeidler1994,Marinacci1999,Grabisch2016},
Shapley proved that there is always a classical probability measure
consistent with the classical IVPM~$\bar{\mu}$ \cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}.
Given any random variable, its expectation value with respect to classical
probability measures consistent with~$\bar{\mu}$ is consistent with
its Choquet integral \cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with respect to $\bar{\mu}$ \cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.

@@

is defined by the maximal length of the intervals, and recovers the

@@

Thus the geometry of a single qubit reduces to transformations among
points on $\Sphere{2}$, which can be parametrized in an infinite
one-parameter family of transformations, one of which is the geodesic
or minimal-length transformation. Explicitly, given two one-qubit
states denoted by points $\Hat{a}$ and $\Hat{b}$ on $\Sphere{2}$,
the shortest rotation carrying $\Hat{a}$ to $\Hat{b}$ is the SLERP
(spherical linear interpolation) \cite{Shoemake:1985:ARQ:325334.325242,wiki:Slerp}
\begin{equation}
S\left(\Hat{a},\Hat{b},t\right)=\Hat{a}\,\frac{\sin((1-t)\omega)}{\sin\omega}\,+\,\Hat{b}\,\frac{\sin(t\omega)}{\sin\omega}\,,\label{slerp.eq}
\end{equation}
where $\Hat{a}\cdot\Hat{b}=\cos\omega$. Figure~\ref{blochSphere.fig}~(b)
illustrates the path traced by a SLERP between two irreducible one-qubit
states on the Bloch sphere. Because states in CQC are defined by infinite
precision real numbers, it is not possible, even in principle, to
make an exact state transition as implied by Fig.~\ref{blochSphere.fig}~(b).
In practice, one has to be content with approximate, typically exponentially
expensive, transitions from state to state.

@@

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event $E\subseteq\Omega$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will actually happen. To maintain
the coherence, $\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
sending exactly the same particle as the previous one, but using a
different beam splitter corresponding to the observable $\mathbf{O}_{1}$
with spectral decomposition $\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$,
where $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$.
Although the sample space $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $
and the probability measure $\mu_{1}:2^{\Omega_{1}}\rightarrow\left[0,1\right]$
defined by this experiment is different from the previous one, these
two experiments may produce the same post-experimental state~$\ket{2}$,
and the probability of the common event $\left\{ \ket{2}\right\} $
is believed to be the same, i.e., $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.
In general, as long as sending the same particle, the probability
of the same event in different experiments should always be the same.
This fact is equivalent to the fact that commuting observables could
be measured simultaneously which is essential to define contextuality
and will be explained in Sec.~\ref{sec:Kochen-Specker}.

>>>>

\usepackage{unicode-math}
\RequirePackage{ifthen}
\@ifpackageloaded{unicode-math}{}{
 \newcommand{\symbf}{\mathbf}
 \newcommand{\symrm}{\mathrm}
 \newcommand{\symcal}{\mathcal}
 \newcommand{\symbb}{\mathbb}
}
\newcommand{\muB}{\ensuremath{\mu^{\symrm{B}}}}
\newcommand{\muC}{\ensuremath{\mu^{\symrm{C}}}}
\newcommand{\muF}{\ensuremath{\mu^{\symrm{F}}}}
\newcommand{\symrmL}{\ensuremath{\symrm{L}}}
\newcommand{\symrmR}{\ensuremath{\symrm{R}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\symrmL{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\symrmR{#1}}}}
\newcommand{\barmuD}{\ensuremath{\bar{\mu}^{\symrm{D}}}}
\newcommand{\events}{\ensuremath{\symcal{E}}}
\newcommand{\eventsC}{\ensuremath{\events_{\symrm{C}}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\symbb{F}_{{#1}}}
\newcommand{\Fp}{\Fpx{p}}
\newcommand{\Fppx}[1]{\Fpx{{#1}^2}}
\newcommand{\Fpp}{\Fppx{p}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\symbf{#1}}}
\newcommand{\mathReal}{\symbb{R}}
\newcommand{\mathComplex}{\symbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\symbf{S}^{#1}}
\newcommand{\R}{\mathReal}
\newcommand{\rme}{\symrm{e}}
\newcommand{\rmi}{\symrm{i}}
\newcommand{\rmd}{\symrm{d}}
\newcommand{\CP}[1]{\mathComplex\symbf{P}^{#1}} % Complex projective space
\newcommand{\DCP}[1]{\symbf{D}\mathComplex\symbf{P}^{#1}} % Discrete complex projective
\def\fh{\mathfrak{h}}
\newcommand{\uf}{U_{\!f}}
\newcommand{\scalarPlus}{+}
\newcommand{\boolt}{\textsf{Bool}} 
\newcommand{\bfalse}{\texttt{\textbf{false}}}
\newcommand{\btrue}{\texttt{\textbf{true}}}
\newcommand{\dotprod}{dot product}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\newcommand{\Psibar}{\overline{\Psi}}
\newcommand{\cardp}[2]{#1 \, \slash\!\! \slash \, #2}
\def\round{\mathop{\rm round}\nolimits}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\Hilb}{\symcal{H}}
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\def\C{{\symbb{C}}}
\newcommand{\ultramodular}{\symcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\frameL}[1][]{\ensuremath{f^{L{#1}}}}
\newcommand{\frameR}[1][]{\ensuremath{f^{R{#1}}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\nb}{\nolinebreak[3] }
\NewDocumentCommand{\jthSubsystem}{m O{} m}{{#1}_{#2}^{#3}}

@@

\item Hao-Chun Lee (李浩君) and Hsien-Ching Kao (高憲慶) for discussing
what I did in the prospective of a general PhD. in Physics and working
later in more computational industry.

@@

Although it is convenient to assume physical quantities is as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick are made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
be an integer multiple of the elementary charge. A prefect square
cannot be draw physically. Even we don't draw it, but merely ask whether
it exists in the space. This question is still beyond our ability
to answer because whether the space itself is ultimately discrete
or continuous is still an unsolved question. Even if a prefect square
exists physically, we may not able to decide whether the ``square''
we discussed is really prefect or almost prefect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision in reality is limited generally three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Hence, when the quantities used to branch the computation closes to
zero, the branch chosen by a BCSS machine cannot reliably predict
the branch chosen by a realistic analog computer. Since two branches
of computations may not have any relations, the follow-up computational
paths may be significantly different, and this difference is unlikely
to be compensated by error analysis techniques easily like the butterfly
effect. Therefore, the computational power of analog computers might
be far weaker than the prediction based on this kind of theoretical
models.%
\begin{comment}
\todo{Maybe extended to consider probabilistic model? }
\end{comment}

@@

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{YcYY}
\toprule 
\addlinespace
Physical quantity & Discrete & Discrete with extreme small units & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational device & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integer & Real number & Real number\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How does the theorical model predict the behavior of the Physical device? & 
Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum model &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and two
possibilities for classical quantities inspires two types of quantum
models. Like its classical counterparts, quantum circuit model, the
most widely used model of quantum computing, manipulates quantities
which are not discrete apparently: the probability amplitudes of quantum
states. To address the similar inconsistency, we first review the
``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}, where the probability amplitudes are in the field of complex numbers.
If the probability amplitudes are assumed to be ultimately discrete,
we choose to replace the field of complex numbers by discrete finite
fields for two reasons. On one hand, CQT is built upon the operations
among the probability amplitudes, and the probability amplitudes have
the required operations since the complex numbers form a field. On
the other hand, based on finite fields, we will define the fraction-like
cardinal probability which has extremely small units when the size
of the field is extremely large. Alternatively, even if the probability
amplitudes might never be found quantized, considering the precision
of quantities could still result in a better model. However, comparing
to the precision of probability amplitudes, it is easier to consider
the precision of their inducing probabilities because the idea of
``imprecise probabilities'' have well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
\todo{Make sure whether realistic quantum computers could branch
their computation based on their probability amplitudes or measured
probabilities? If they can only branch on their measured probabilities,
it will provide another reason why we want to study the precision
of the inducing probabilities\ldots{} }This idea will be extended
to quantum domain as quantum interval-valued probability measures
(QIVPM).

@@

Finally, in Section~\ref{discretequantumcomputingII}, we apply our
discrete quantum theory to the study of two representative algorithms,
the deterministic Deutsch-Jozsa algorithm and the probabilistic Grover
algorithm~\cite{NCbook,Mermin}. The first algorithm highlights the
role played by the size of the field $p$ in determining the actual
resources required for computation as the number of input bits $n$
increases, a concept nonexistent in conventional quantum computing.
The second algorithm highlights, in addition, the dependence of the
precision of measurement (via cardinal probabilities) on the size
of the field, another nonexistent concept in conventional quantum
computing.
\end{comment}
\begin{comment}
Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\end{comment}
Since many important quantum algorithms are probabilistic, we want
to find a notion of probability for quantum theories over finite fields.
The cardinal probability defined in Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
restricts states in some local regions of the vector spaces, where
a notion of inner product and probability could be recovered. In Sec.~\ref{discretequantumcomputingII},
this discrete quantum theory is applied on two representative algorithms,
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\begin{comment}
When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can be deduced from a set of abstract definitions and axioms
according to Gleason's theorem. Although we might hope to deduce a
discrete Born rule directly from a similar set of definitions and
axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability and extend it with the conventional quantum theory.
\end{comment}
Despite the great success of cardinal probabilities, it is difficult
to define their arithmetic operations and expectation values. Since
they could be defined easily with real-valued probabilities in CQT,
we attempt to consider real-valued quantum probability measures over
finite fields (QPMFF) in Sec.~\ref{sec:Toward-IVPM}. In contrast
to CQT, our Gleason-like conditions over finite fields cannot deduce
a discrete Born rule in general. This impossible result suggests us
that the discrete and finite state spaces might not be compatible
with infinitely precise real-valued probability. Thus, no matter whether
the probability amplitudes are ultimately discrete or continuous,
we should study imprecise quantum probabilities formulated by quantum
interval-valued probability measures.

\begin{comment}
In the classical setting, there are several proposals for ``imprecise
probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
Although these proposals differ in some details, they all share the
fact that the probability $\bar{\mu}(E)$ of an event~$E$ is generalized
from a single \emph{real number} to an \emph{interval}~$[l,r]$,
where~$l$ intuitively corresponds to the strength of evidence for
the event~$E$ and~$1-r$ corresponds to the strength of evidence
against the same event. Given a sample space~$\Omega$ and a set
of intervals~$\mathscr{I}$, like a classical probability measure
$\mu:2^{\Omega}\rightarrow\left[0,1\right]$, a classical interval-valued
probability measure (IVPM) $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
needs to satisfy some coherent axioms. By satisfying the convexity
axiom \cite{Shapley1971,GilboaSchmeidler1994,Marinacci1999,Grabisch2016},
Shapley proved that there is always a classical probability measure
consistent with the classical IVPM~$\bar{\mu}$ \cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}.
Given any random variable, its expectation value with respect to classical
probability measures consistent with~$\bar{\mu}$ is consistent with
its Choquet integral \cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with respect to $\bar{\mu}$ \cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}.

@@

is defined by the maximal length of the intervals and recovers the

@@

Thus the geometry of a single qubit reduces to transformations among
points on $\Sphere{2}$, which can be parametrized in an infinite
one-parameter family of transformations, one of which is the geodesic
or minimal-length transformation. Explicitly, given two one-qubit
states denoted by points $\Hat{a}$ and $\Hat{b}$ on $\Sphere{2}$,
the shortest rotation carrying $\Hat{a}$ to $\Hat{b}$ is the SLERP
(spherical linear interpolation) \cite{Shoemake:1985:ARQ:325334.325242,wiki:Slerp}
\begin{equation}
S\left(\Hat{a},\Hat{b},t\right)=\Hat{a}\,\frac{\sin((1-t)\omega)}{\sin\omega}\,+\,\Hat{b}\,\frac{\sin(t\omega)}{\sin\omega}\,,\label{slerp.eq}
\end{equation}
where $\Hat{a}\cdot\Hat{b}=\cos\omega$. Figure~\ref{blochSphere.fig}~(b)
illustrates the path traced by a SLERP between two irreducible one-qubit
states on the Bloch sphere. Because states in CQC are defined by infinite
precision real numbers, it is not possible, even in principle, to
make an exact state transition as implied by Fig.~\ref{blochSphere.fig}~(b).
In practice, one must be content with approximate, typically exponentially
expensive, transitions from state to state.

@@

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event $E\subseteq\Omega$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will happen. To maintain the coherence,
$\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
sending exactly the same particle as the previous one but using a
different beam splitter corresponding to the observable $\mathbf{O}_{1}$
with spectral decomposition $\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$,
where $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$.
Although the sample space $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $
and the probability measure $\mu_{1}:2^{\Omega_{1}}\rightarrow\left[0,1\right]$
defined by this experiment is different from the previous one, these
two experiments may produce the same post-experimental state~$\ket{2}$,
and the probability of the common event $\left\{ \ket{2}\right\} $
is believed to be the same, i.e., $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.
In general, as long as sending the same particle, the probability
of the same event in different experiments should always be the same.
This fact is equivalent to the fact that commuting observables could
be measured simultaneously which is essential to define contextuality
and will be explained in Sec.~\ref{sec:Kochen-Specker}.
